test

### 2023-03-08 test.Rmd
Update test.Rmd
Removed reading in chunks. This will be code only

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Read in data

```{r}
# Read in soundscape data, round to nearest 20 min
BB_08_csv <- read_csv("Data/CCES_08_1Hz_1s_BB_2min.csv")
BB_08 <- BB_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
BB_08 <- BB_08 %>%
  mutate(dateTime = round_date(ymd_hms(BB_08$dateTime),"20 minutes"))
```

### Read in Tracks
```{r}
# Read in tracks, filter out lost buoys
tracks_csv <- read_csv("Data/AllTracks.csv")
tracks <- tracks_csv %>%
  dplyr::filter(station == "4"| station == "7"| station =="8" |station == "10"| 
                station == "12"|station == "13"| station == "14"| station == "16"|
                station == "17"|station == "18"| station == "19"| station == "20"| 
                station == "21"|station == "22"| station == "23")

# Filter by CCES_08 buoy
tracks_08 <- tracks %>% dplyr::filter(station == "8")

# Round CCES_08 to nearest 20 min
tracks_08 <- tracks_08 %>% 
  mutate(dateTime = ceiling_date(ymd_hms(tracks_08$dateTime), "20 minutes"))

@@

### Read in OL
```{r}
# Read in OL data, round to nearest 20 min
OL_08_csv <- read_csv("Data/CCES_08_1Hz_1s_OL_2min.csv")
OL_08 <- OL_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
OL_08 <- OL_08 %>%
  mutate(dateTime = round_date(ymd_hms(OL_08$dateTime),"20 minutes"))
OL_08

@@

Joining tables
```{r}
# Join tables
track_BB_08 <- left_join(tracks_08_rnd, BB_08, by = "dateTime")
track_BB_OL_08 <- left_join(track_BB_08, OL_08, by="dateTime")

write_csv(track_BB_OL_08, "Data/track_bb_OL_08.csv")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sf)
library(rerddap)
library(terra)
library(marmap)
library(rnaturalearth)
library(mapproj)
library(raster)
library("rerddapXtracto")
library(PAMpal)
library(maptiles)
library(tmap)
library(spData)
library(tmaptools)
library(grid)
library(ggspatial)
```

## Read in soundscape data, NEED TO MAKE THIS A FUNCTION

```{r}
# Round to nearest 20 minutes, create dateTime data type
# Drift 8: Broadband
BB_08_csv <- read_csv("Data/CCES_08_1Hz_1s_BB_2min.csv", show_col_types = FALSE)
BB_08 <- BB_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
BB_08 <- BB_08 %>%
  mutate(dateTime = round_date(ymd_hms(BB_08$dateTime),"20 minutes"))

# Drift 8: Octave Level
OL_08_csv <- read_csv("Data/CCES_08_1Hz_1s_OL_2min.csv",show_col_types = FALSE)
OL_08 <- OL_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
OL_08 <- OL_08 %>%
  mutate(dateTime = round_date(ymd_hms(OL_08$dateTime),"20 minutes"))

# Drift 8: TOL
TOL_08_csv<- read_csv("data/CCES_08_1Hz_1s_TOL_2min.csv", show_col_types = FALSE)
TOL_08 <- TOL_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
TOL_08 <- TOL_08 %>%
  mutate(dateTime = round_date(ymd_hms(TOL_08$dateTime),"20 minutes"))

# Drift 10: Broadband
BB_10_csv <- read_csv("Data/CCES_10_1Hz_1s_BB_2min.csv", show_col_types = FALSE)
BB_10 <- BB_10_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
BB_10 <- BB_10 %>%
  mutate(dateTime = round_date(ymd_hms(BB_10$dateTime),"20 minutes"))

# Drift 10: TOL
TOL_10_csv<- read_csv("data/CCES_10_1Hz_1s_TOL_2min.csv", show_col_types = FALSE)
TOL_10 <- TOL_10_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
TOL_10 <- TOL_10 %>%
  mutate(dateTime = round_date(ymd_hms(TOL_10$dateTime),"20 minutes"))

# Drift 10: Octave level
OL_10_csv <- read_csv("Data/CCES_10_1Hz_1s_OL_2min.csv", show_col_types = FALSE)
OL_10 <- OL_10_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
OL_10 <- OL_10 %>%
  mutate(dateTime = round_date(ymd_hms(OL_10$dateTime),"20 minutes"))

@@

### Read in Tracks

```{r}
# Read in tracks, filter out lost buoys
tracks <- read_csv("Data/AllTracks.csv", show_col_types = FALSE) %>%
  dplyr::filter(station == "4"| station == "7"| station =="8" |station == "10"|          # filter out lost buoys
                station == "12"|station == "13"| station == "14"| station == "16"|
                station == "17"|station == "18"| station == "19"| station == "20"| 
                station == "21"|station == "22"| station == "23") %>%
  mutate(dateTime = round_date(ymd_hms(tracks$dateTime), "20 minutes"))                 
  # round to nearest 20 minutes

# Filter by buoy 8 and 10
tracks_08 <- tracks %>% dplyr::filter(station == "8")
tracks_10 <- tracks %>% dplyr::filter(station == "10")

# Tracks on a map
tracks_sf <- tracks %>% # make tracks an sf object
  st_as_sf(coords= c("long", "lat"), crs=4326, remove=F)

@@

### Read in whale detections

```{r}
# All whale detections
whales <- read_csv("Data/EventInfo.csv", show_col_types = FALSE) %>%
  mutate(lat = as.numeric(Latitude)) %>%
  mutate(long = as.numeric(Longitude)) %>%
  mutate(bestNumber = as.numeric(bestNumber)) %>%   #convert nClicks from character to numeric
  mutate(nClicks = as.numeric(nClicks)) %>%      # convert nClicks from character to numeric
  mutate(StartTime = mdy_hm(StartTime)) %>%      #convert time from character to datetime
  mutate(EndTime = mdy_hm(EndTime)) %>%
  drop_na(...1)%>%                                 # remove NAs in column labeled "...1"
  dplyr::select(-Project)

BWsf <- whales %>% st_as_sf(coords = c("long","lat"), crs=4326)   # whales as sf

# Drift 10 whales as sf
whales10_sf <- BWsf %>%                            
  dplyr::filter(Deployment == "10")

whales08_sf <- BWsf %>%                            
  dplyr::filter(Deployment == "8")

# Drift 10 whales
whales10 <- whales %>%
  dplyr::filter(Deployment == "10")

# Drift 8 whales
whales8 <- whales %>%
  dplyr::filter(Deployment == "8")
```

### Joining tracks with soundscape metrics with whale detections

```{r}
# Drift 08
tracks_SS_08 <- left_join(tracks_08, BB_08, by = "dateTime") %>%   # join tracks with BB 08
  left_join(., TOL_08, by = "dateTime") %>%                        # join TOL
  left_join(., whales8, by = "long") %>%                           # join whales 
  filter(spotID == "0-2571609") %>%                                # filter by one SPOT
  mutate(BWpresence = if_else(is.na(nClicks), 0, 1)) %>%           # Add column to specify BW presence (1) or absence (0)
  dplyr::select(dateTime, lat.x, long, station, `BB_20-24000`, TOL_63, 
                TOL_125, TOL_2000, TOL_5000, TOL_20000, species, BWpresence)  # keep necessary columns

#Drift 10
tracks_SS_10 <- left_join(tracks_10, BB_10, by = "dateTime") %>%    # join tracks with BB 10
  left_join(., TOL_10, by = "dateTime") %>%                         # join tracks with TOL 10 
  left_join(., whales10, by = "long") %>%                           # join whales 
  filter(spotID == "0-2572490") %>%                                 # filter by one SPOT
  mutate(BWpresence = if_else(is.na(nClicks), 0, 1)) %>%           # Add column to specify BW presence (1) or absence (0)
  dplyr::select(dateTime, lat.x, long, station, `BB_20-24000`, TOL_63, 
                TOL_125, TOL_2000, TOL_5000, TOL_20000, species, BWpresence)  # keep necessary columns
```

## AIS

For 10-01-18 to 10-10-18

```{r}
rawAIS <- read_csv("Data/AIS_100118_101018.csv", show_col_types = FALSE) # Vessel data from 10-01-18 to 10-10-18
drift <- track_BB_OL_10

# Filter by vessels over 100m in length, remove unnecessary columns
AIS <- rawAIS %>%
  filter(Length >= 100) %>%
  dplyr::select(-SOG,-COG, -CallSign, -IMO, -Cargo, -Draft, -Status)

# Currently unused
bbox <- st_bbox(c(xmin = min(drift$lat-0.7), 
            xmax = max(drift$lat + 0.7), 
            ymax = max(drift$long + 0.7), 
            ymin = min(drift$long-0.7)), crs = st_crs(4326))

AIS_sf <- AIS %>%
  st_as_sf(coords = c("LON", "LAT"), crs=4326, remove=F) #%>%
  #st_crop(y = bbox)

AIS_sf <- AIS_sf %>% 
  filter(BaseDateTime > "2018-10-05 00:00:00")  # Shorten time range to 10/05/2018 to 10/10/2018
```

#### AIS

For 08-10-18 to 08-17-18

```{r}
rawAIS <- read_csv("Data/AIS_081018_081718.csv", show_col_types = FALSE) # Vessel data from 08-10-18 to 08-17-18

augAIS <- rawAIS %>% filter(Length >= 100) %>%
  dplyr::select(-SOG,-COG, -CallSign, -IMO, -Cargo, -Draft, -Status)

### Exploratory analysis

#### Not important: sound by latitude and dateTime

```{r}
# Does sound vary with latitude?
lm(formula = OL_63 ~ lat, data = track_BB_OL_08)
fit10 <- lm(formula = lat ~ OL_63, data = track_BB_OL_10)
plot(fit10)
ggplot(track_BB_OL_08, aes(x = lat, y = `BB_20-24000`)) + scale_y_continuous() +
   geom_point() + geom_smooth()

# Does sound vary with date/time?
summary(lm(formula = `BB_20-24000`~ dateTime, data = track_BB_OL_10))

ggplot(track_BB_OL_08, aes(x = dateTime, y = `BB_20-24000`)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Buoy 08 - Sound levels by date")

ggplot(track_BB_OL_10, aes(x = dateTime, y = `BB_20-24000`)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Buoy 10 - Sound levels by date")

plot(lm(formula = `BB_20-24000`~ dateTime, data = track_BB_OL_08))

ggplot() +
  geom_point(data = track_BB_OL_08, aes(x = dateTime, y = `BB_20-24000`, color = "red")) +
  geom_point(data = track_BB_OL_10, aes(x = dateTime, y = `BB_20-24000`, color = "blue"))
```

#### Not important: tracks as sf and basemaps

```{r}
library(tmap); library(maptiles); library(sf)
tmap_mode("plot")
trackssf <- st_as_sf(tracks, coords = c("long", "lat"), crs=4326)
tracks08sf <- st_as_sf(track_BB_OL_08, coords = c("long", "lat"), crs=4326)
tracks10sf <- st_as_sf(track_BB_OL_10, coords = c("long", "lat"), crs=4326)
trk_TOL_08 <- st_as_sf(tracks_SS_08, coords = c("long", "lat.x"), crs=4326)
trk_TOL_10 <- st_as_sf(tracks_SS_10, coords = c("long", "lat.x"), crs=4326)
trackBase <- get_tiles(whales_sf, provider = "Esri.WorldShadedRelief")
allTrackbase <- get_tiles(trackssf, provider = "Stamen.TonerLite", zoom = 5)
```

### Sea Surface Temp with rerddap, Drift 8 only
```{r}
SSTInfo <- info('jplMURSST41')  # Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Daily 
murSST <- griddap(sstInfo, latitude = c(35, 39), longitude = c(-122, -130), time = c("2018-08-16", "2018-10-01"), fields = 'analysed_sst') # drift8 dates only
SST <- murSST$data

# Raster of SST
ggplot() + 
  geom_tile(data = SST, aes(x = longitude, y = latitude, fill = analysed_sst)) +
  scale_fill_gradient(low = "darkblue", high = "yellow") +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        legend.position = "none")
```

### SST from NOAA CoastWatch : https://coastwatch.pfeg.noaa.gov/projects/r/xyt-matchup.html
```{r}
xcoord <- tracks_08$long
ycoord <- tracks_08$lat
tcoord <- tracks_08$dateTime
xlen <- 0.2
ylen <- 0.2
parameter <- "analysed_sst"
dataset <- 'jplMURSST41'
#dataInfo <- rerddap::info(dataset, url= "https://coastwatch.pfeg.noaa.gov/erddap/")

#swchl <- rxtracto(dataInfo, 
#                  parameter=parameter, 
#                  xcoord=xcoord, ycoord=ycoord, 
#                  tcoord=tcoord, xlen=xlen, ylen=ylen)

#plotTrack(swchl, xcoord, ycoord, tcoord, plotColor = 'thermal')
```

### SST from PAMPal
```{r}
# Downloads data from servers
# tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'jplMURSST41', fileName = "analysedSST")

# from file 
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'analysedSST.nc')
tracks_SS_08
```

### Distance to continental slope

```{r}
# Increase cell size by 30x to be visible
bathy <- rast("data/gebco_bathy.tif")
agg <- terra::aggregate(bathy, 20)
plot(agg)

# Rasterize track with agg as the template
#track_rast <- rasterize(tracks08sf, agg, field = "BB_20-24000")
#TOL_rast <- rasterize(trk_TOL_08, agg, field = "TOL_125")
sf08 <- st_as_sf(tracks_SS_08, coords = c("long", "lat.x"), crs=4326, remove=F)
sf10 <- st_as_sf(tracks_SS_10, coords = c("long", "lat.x"), crs=4326, remove=F)
trkraster08 <- rasterize(vect(sf08), agg)
trkraster10 <- rasterize(vect(sf10), agg)

bathy_df <- as.data.frame(bathy, xy = TRUE)

# Map the coastline and cont. slope
cont <- st_as_sf(as.contour(bathy))
coast <- cont %>% filter(level == 0)
cont_slope <- cont %>% filter(level == -2000)
cont_slope_rast <- rasterize(cont_slope, agg, field = "level")
plot(cont_slope_rast, col = "red")
cont_slope_rast <- rasterize(cont_slope, agg)
plot(cont_slope_rast, col = "red")

ggplot() + 
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y, 
                            alpha = 0.8,
                            fill = `gebco_bathy`))+
  geom_sf(data = cont, col="black") + 
  geom_sf(data = coast, col = "red") +
  geom_sf(data = cont_slope, col = "purple") +
  geom_sf(data = tracks08sf, mapping = aes(col= `BB_20-24000`)) +
  geom_sf(data = tracks10sf, mapping = aes(col= `BB_20-24000`)) +
  scale_color_gradient(low="green", high="red") 
```

### Distance to continental slope

```{r}
# Drift 8
slopedist <- terra::distance(trkraster08, vect(cont_slope))     # measure distance from track to slope
extdist <- terra::extract(slopedist, vect(sf08))       # extract this distance into a df
tracks_SS_08 <- cbind(sf08,extdist) %>%                     # bind to main df
  dplyr::select(-ID) %>%
  rename("dist_slope" = gebco_bathy)

#Drift 10
slopedist <- terra::distance(trkraster10, vect(cont_slope))     # measure distance from track to slope
extdist <- terra::extract(slopedist, vect(sf10))       # extract this distance into a df
tracks_SS_10 <- cbind(sf10,extdist) %>%                     # bind to main df
  dplyr::select(-ID) %>%
  rename("dist_slope" = gebco_bathy)
```

#### Plot distance to continental slope

```{r}
# By color
ggplot() +
  geom_sf(data = pts08, 
          mapping = aes(col = `Distance to Continental Slope`)) +
  geom_sf(data = pts10, 
          mapping = aes(col = `Distance to Continental Slope`)) +
  geom_sf(data = cont_slope)

# Plot distance by OL/TOL 
ggplot(data = pts08, aes(x = `Distance to Continental Slope`, y = `OL_125`)) + geom_point()
ggplot(data = TOL08, aes(x = `Distance to Continental Slope`, y = `TOL_125`)) + geom_point() + geom_smooth()
   # what are the units for distance? Meters

ggplot() +
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y,
                            fill = `gebco_bathy`)) +
  geom_sf(data = TOL08, 
          mapping = aes(col = `Distance to Continental Slope`)) +
  geom_sf(data = cont_slope) + scale_fill_continuous(low = "lightblue", high = "darkblue")

summary(lm(formula = BB_20.24000 ~ `Distance to Continental Slope`, data = pts08))

# PLOT distance by TOL with whales
ggplot(data = TOL_whale_08, aes(x = `Distance to Continental Slope`, y = `TOL_125`)) + 
  geom_point() + 
  geom_smooth(color = "green") +
  theme_classic()

# SAME PLOT with size as whale detections
ggplot(data = TOL_whale_08, aes(x = `Distance to Continental Slope`, y = `TOL_125`)) + 
  geom_point(aes(size = minNumber)) + 
  geom_smooth() +
  theme_classic()
```

### Third octave levels histograms

```{r}
# Drift 8 - all TOL bands, facet wrap
pivot <- tracks_SS_08 %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: Drift 8") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.85,0.2)) +
  facet_wrap(~TOL)

# Drift 10 - all TOL bands, facet grid
pivot2 <- tracks_SS_10 %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot2, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 5, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: Drift 10") +
  scale_x_continuous(name = "Sound Intensity (dB)", breaks=seq(50,110,20)) +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.76)) +
  facet_grid(~TOL)

# Drift 8 - TOL bands 125 and 2000 only
pivot3 <- tracks_SS_08 %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz")

ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("DASBR 8: Distribution of Third-Octave Level (TOL) Intensities") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))

# Drift 10 - TOL bands 125 and 2000 only
pivot4 <- tracks_SS_10 %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot4, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  #ggtitle("DASBR 10: Distribution of Third-Octave Level (TOL) Intensities") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))

# TOL linear model plots
#plot(lm(formula = `Distance to Continental Slope` ~ bestNumber, data = TOL_whale_08))

#ggplot(TOL_whale_08) +
#  geom_point(aes(x = TOL_125, y = nClicks), color = "blue") #+
#  geom_point(aes(x = TOL_63, y =  nClicks), color = "hotpink") +
#  geom_point(aes(x = TOL_5000, y =  nClicks), color = "limegreen") +
#  geom_point(aes(x = TOL_2000, y =  nClicks), color = "yellow")

### Does hour of day affect sound levels?

```{r}
# CCES_08
hr_08 <- track_BB_OL_08 %>%
  mutate(Hr = hour(track_BB_OL_08$dateTime)) %>%
  group_by(Hr) %>%
  summarize(meanBB = mean(`BB_20-24000`)) %>%
  rename(`Buoy 08 Broadband` = `meanBB`)
ggplot(data = hr_08, mapping = aes(x = Hr, y = `Buoy 08 Broadband`)) + 
  geom_point() + geom_smooth()

#CCES_10
hr_10 <- track_BB_OL_10 %>%
  mutate(Hr = hour(track_BB_OL_10$dateTime)) %>%
  group_by(Hr) %>%
  summarize(meanBB = mean(`BB_20-24000`)) %>%
  rename(`Buoy 10 Broadband` = `meanBB`)
ggplot(data = hr_10, mapping = aes(x = Hr, y = `Buoy 10 Broadband`)) + 
  geom_point() + geom_smooth()

# Join and graph
hr_join <- left_join(hr_08, hr_10, by = "Hr")
hr_piv <- hr_join %>%
  pivot_longer(col=c(`Buoy 08 Broadband`,`Buoy 10 Broadband`), names_to = "buoys", values_to = "BB")

ggplot(data = hr_piv, mapping = aes(Hr, BB, col=buoys)) +
  geom_point() + geom_smooth()
```

### Are sound levels affected by species composition or number of clicks?

```{r}
# Plot
library(tmap); library(maptiles)
tmap_mode("plot")
oceanBase <- get_tiles(whales_sf, provider="Esri.NatGeoWorldMap", crop = TRUE)
tm_shape(oceanBase) + tm_rgb() +
  tm_shape(whales_sf) + 
  tm_symbols(col="species", size = "nClicks") +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"))

# CCES 08
BB_whales08 <- st_join(pts08, whales_sf) %>%
  filter(!is.na(`species`))

summary(glm(`BB_20.24000` ~ species + `Distance to Continental Slope` + nClicks, data = BB_whales08))

# CCES 10
BB_whales10 <- st_join(pts10, whales_sf) %>%
  filter(!is.na(`species`))

summary(glm(`BB_20.24000` ~ species + `Distance to Continental Slope` + nClicks, data = BB_whales10))

#Raster base map
tmap_mode("plot")
oceanBase <- get_tiles(BB_whales10, provider="Esri.NatGeoWorldMap")
tm_shape(bathy) + tm_raster() +
  tm_shape(BB_whales10) + 
  tm_symbols(col="species", size = "nClicks", style = "jenks") +
  tm_layout(legend.width = 0.5)

### Explore environmental variables

#### Bathymetry

```{r}
library(raster)
bgo <- getNOAA.bathy(lon1 = -121, lon2 = -131,
                     lat1 = 34, lat2= 40, resolution = 1)  # plot bounding box around drift 8 and 10
# As raster
bgo_rast <- marmap::as.raster(bgo)

# Get depth of whale detections in drift 8
depth08 <- get.depth(bgo, TOL_whale_08[,4:3], locator = FALSE)

# Get depth of tracks
trkdepth08 <- get.depth(bgo, TOL_whale_08[,4:3], locator = FALSE)
trkjoin <- left_join(TOL_whale_08, trkdepth08, by = c("long"= "lon")) %>%
  mutate(depth = depth*-1)%>%     # remove the negative sign
  mutate(nClicks = coalesce(nClicks, 0))   # replace NAs with zeros

#Plot
bplot + geom_point(data = depth08, aes(x = lon, y = lat, color = depth, size = 2)) +
  scale_color_gradient(low = "red", high = "beige") +
  ggtitle("Depth of Drfit 8 whale detections")

# Join depth08 with TOL_whale_08 on long
join <- left_join(TOL_whale_08, depth08, by = c("long" = "lon"))

ggplot(trkjoin) + geom_point(aes(x = depth, y = nClicks))

summary(glm(formula = nClicks ~ TOL_125 + `Distance to Continental Slope` + depth + TOL_63 + TOL_2000 + TOL_5000, data = trkjoin, family = poisson))
```

#### Bathymetric slope

```{r}
# Slope
bgo_slope <- terrain(bgo_rast, v = "slope")
plot(bgo_slope, axes = FALSE)

# Extract slope values at points along drift 8
slope <- raster::extract(bgo_slope, tracks_SS_08[,4:3])
tracks_SS_08 <- cbind(tracks_SS_08, slope)

#ggplot() +
#  geom_raster(data = bgo_slope, mapping = aes(x=x, y=y, fill=slope)) + 
#  geom_point(data = depth08, aes(x = lon, y = lat, color = depth, size = 2))

# Roughness (different from rugosity?)
plot(terrain(bgo_rast, opt = "roughness"))    # what are the units on this?
```

#### Distance from coast - PAMPal

```{r}
# PAMpal function
tracks_SS_08 <- as.data.frame(tracks_SS_08) %>% 
  mutate(Latitude = lat.x, 
         Longitude = long, 
         UTC = as.POSIXct(dateTime, format='%Y-%m-%d %H:%M:%S', tz='UTC'))

dist2shore <- erddapToEdinfo(dataset='dist2coast_1deg',
                             baseurl='https://pae-paha.pacioos.hawaii.edu/erddap/')
# Downloads nc file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc=dist2shore, fileName='Dist2Shore.nc')

tracks_SS_08 <- matchEnvData(tracks_SS_08, nc='Dist2Shore.nc')  
```

#### Chlorophyll-a: NOAA Coastwatch

```{r}
# Code chunk taken from NOAA CoastWatch : https://coastwatch.pfeg.noaa.gov/projects/r/xyt-matchup.html
xcoord <- tracks_08_rnd$long
ycoord <- tracks_08_rnd$lat
tcoord <- tracks_08_rnd$dateTime
xlen <- 0.2
ylen <- 0.2
parameter <- "chlorophyll"
dataset <- 'erdMH1chlamday'
dataInfo <- rerddap::info(dataset, url= "https://coastwatch.pfeg.noaa.gov/erddap/")

swchl <- rxtracto(dataInfo, 
                  parameter=parameter, 
                  xcoord=xcoord, ycoord=ycoord,
                  tcoord=tcoord, xlen=xlen, ylen=ylen)

plotTrack(swchl, xcoord, ycoord, tcoord, plotColor = 'algae')

### Chlorophyll from PAMPal
```{r}
# Downloads from server
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdMH1chlamday', fileName='chlorophyll.nc')

# Pulls from downloaded file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'chlorophyll.nc')
tracks_SS_08
```

#### Sea surface height anomalies - NOAA Coastwatch
```{r}
# Code chunk taken from NOAA CoastWatch : https://coastwatch.pfeg.noaa.gov/projects/r/xyt-matchup.html
xcoord <- tracks_08_rnd$long
ycoord <- tracks_08_rnd$lat
tcoord <- tracks_08_rnd$dateTime
xlen <- 0.2
ylen <- 0.2
parameter <- 'sla'
dataset <- 'nesdisSSH1day'
dataInfo <- rerddap::info(dataset, url= "https://coastwatch.pfeg.noaa.gov/erddap/")

swssh <- rxtracto(dataInfo, 
                  parameter=parameter, 
                  xcoord=xcoord, ycoord=ycoord, 
                  tcoord=tcoord, xlen=xlen, ylen=ylen)

plotTrack(swssh, xcoord, ycoord, tcoord, plotColor = 'deep')
```

#### Sea Surface Height Anomalies from PAMPal
```{r}
# Downloaded from server
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'nesdisSSH1day', fileName='sshanomalies.nc')

# Pulled from downloaded nc file
tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'sshanomalies.nc')
tracks_SS_08
```

#### HyCom Data - MLDepth, MLTemp, TTemp, TDepth, Salinity
```{r}
hycomData <- matchEnvData(tracks_SS_08, nc=PAMmisc::hycomList, raw=TRUE, depth = 0:1000, fileName = 'hycomData')

### Plot maps for Powerpoint images

Map of whole study area
```{r}
# Establish bounding box
bbox <- st_bbox(c(xmin = -110, 
            xmax = -128, 
            ymax = 44, 
            ymin = 30), crs = st_crs(4326))

#Establish basemap
allTrackbase <- get_tiles(bbox, provider = "Esri.WorldGrayCanvas", zoom = 5)

# Establish inset map
us_region <- st_bbox(c(xmin = -90,
            xmax = -120, 
            ymax = 48, 
            ymin = 25), crs = st_crs(4326))
usa <- get_tiles(us_region, provider = "Esri.WorldGrayCanvas", zoom = 4)

sg <- bb_poly(bb(tracks_sf))
inset <- tm_shape(usa) + tm_rgb() + tm_shape(sg) + tm_borders(col = "red")

#Map all tracks
main <- tm_shape(allTrackbase) + tm_rgb() +
  tm_shape(trackssf) +
  tm_grid(alpha = 0.6, col = "grey") +
  tm_dots(col = "darkblue", size = 0.001, legend.show = TRUE) +
  tm_scale_bar(width = 0.3, text.size = 3, position = c(0.2, 0.03)) +
  tm_compass(type = "arrow", position = c(0.05, 0.05), color.light = "white") +
  tm_add_legend('symbol',
                type = "line",
                col = "darkblue",
                labels = "Drifting buoy tracks",
                size = 1,
                lwd = 2)

vp <- viewport(x=0.66, y=0.8, width = 0.26, height=0.4)
main
print(inset,vp=vp)

# Map all tracks - ggplot
basemap2 <- as.data.frame(allTrackbase, xy = TRUE)
#ggplot() +
#  geom_tile(data = basemap, aes(x=x, y=y, fill = "basemap")) +
#  geom_sf(data = trackssf, mapping = aes(), color = "red")

# Map tracks 08 and 10 only
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(tracks08sf) + tm_dots(col = "red", 
                                 legend.show = TRUE) +
  tm_shape(tracks10sf) + tm_dots(col = "green")

# Bounding box of drift tracks on map
lon2 <- -129.04666
lon1 <- -115.78247
lat1 <- 28.21790
lat2 <- 45.09226

basemap3 <- ne_countries(scale = "medium", type = "map_units", returnclass = "sf")

# Drift tracks
ggplot(data = basemap3) +
  geom_sf(fill = "antiquewhite") +
  coord_sf(xlim = c(lon2 - 1, lon1 + 1), ylim = c(lat1 - 1, lat2 + 1), expand = FALSE) +
  theme(panel.grid.major = element_blank()) +
  geom_path(data = tracks_sf, mapping = aes(x=long,y=lat, group = factor(station)), color = "deeppink") +
  theme(panel.background = element_rect(fill = "aliceblue"), 
        axis.text = element_blank(), axis.ticks = element_blank(), 
        axis.title = element_blank())
```

Map of soundscapes
```{r}
# CCES_08 + CCES_10 Broadband
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(tracks08sf) + 
  tm_symbols(col = "BB_20-24000", 
             border.lwd = NA, 
             size = 0.4, 
             sizes.legend = 0.7, 
             n = 10) +
  tm_shape(tracks10sf) + 
  tm_symbols(col = "BB_20-24000",
             border.lwd = NA, 
             size = 0.4,
             n = 10) +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"),
            legend.bg.color = "lightblue")

#CCES_08 Octave Level
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(tracks08sf) + 
  tm_symbols(col = "OL_63", 
             border.lwd = NA, 
             size = 0.5, 
             n = 10) + 
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top")) + 
  tm_legend()

ggplot() + 
  geom_sf(data = trk_TOL_08, aes(color = TOL_125, size = TOL_125), show.legend = FALSE) +
  geom_sf(data = trk_TOL_10, aes(color = TOL_125, size = TOL_125)) +
  scale_x_continuous(limits = c(-130,-122)) + 
  scale_y_continuous(limits = c(35,39)) +
  scale_color_gradient(low = "navy", high = "yellow")+
  scale_size(range = c(1,12)) +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none")
```

Map of whale detections
```{r}
# Whale detections
ggplot() +
  geom_point(data = whales8, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  geom_point(data = whales10, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  scale_x_continuous(limits = c(-130,-122)) + 
  scale_y_continuous(limits = c(35,39)) +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank()
        )
#ggsave("BWdetect.png")
